MLExplain

### 2018-05-24 paper.tex
adding details to JSExplain, WIP

\begin{abstract}
  MLExplain is a step-by-step visual interpreter for OCaml that enables the
  user to inspect both their program's state and the interpreter's state
  itself. This interpreter aims to provide the user with a better understanding
  of the semantics of OCaml semantics.
\end{abstract}

@@

The project \emph{JSExplain} \cite{chargueraud:hal-01745792} aims
to describe a JavaScript program's execution by showing every step of an
interpreter which behavior is the closest to JavaScript specification. This
paper shows how we adapted JSExplain to OCaml.

Unlike JavaScript, OCaml has no official specification. However, OCaml code
execution is fairly simple because the amount of language constructions is
small. We have written an interpreter for OCaml's typed abstract syntax tree
(AST). This AST is close to source code and need little transformations but
supplies us with resolved names -- which is useful for module and signatures
inclusion and absolutely mandatory for features like named parameters in
functions.  The semantic we give to OCaml is higher level than the one
described in \emph{ZINC} \cite{Leroy-ZINC} virtual machine which is at the same
level as object code.

\section{The visual interpreter JSExplain}

An interpreter is a program that takes code written in a specific language as
input and execute it. In order to achieve its goal, the interpreter does
several analyses ; here are the most common : lexical analysis, parsing and
typing.

We call a \emph{visual interpreter} an interpreter that is able to display the
program's state, the current location in source and local-variable values at
each step of the execution as well as the state of the interpreter itself.
JSExplain is a visual interpreter for JavaScript. The aim of this tool is to
enable a user to better understand JavaScript's semantics. JSExplain's
interface highlights the expression of the program currently evaluated as well
as the instruction of the interpreter that is executed.

\section{JSExplain functioning}

>>>>

\begin{abstract}
  MLExplain is a step-by-step interpreter for OCaml that enables the
  user to inspect both their program's state and the interpreter's state
  itself. This interpreter aims to provide the user with a better understanding
  of the semantics of OCaml semantics.
\end{abstract}

@@

The goal of the \emph{JSExplain} project \cite{chargueraud:hal-01745792} is to
describe the execution of a JavaScript program by showing every step of an
interpreter whose behavior is very closest to the specification of JavaScript.
In this paper, we show how we adapted JSExplain to OCaml.

Unlike JavaScript, OCaml has no official specification. However, OCaml code
execution is fairly simple because the amount of language constructions is
small. We have written an interpreter for OCaml's typed abstract syntax tree
(AST). This AST is close to the source code and need little transformations but
supplies us with resolved names -- which is useful for module and signatures
inclusion and absolutely mandatory for features like named parameters and
optional parameters in functions. The semantic we give to OCaml is higher level
than the one described in \emph{ZINC} \cite{Leroy-ZINC} virtual machine which is
at the same level as bytecode.

\section{JSExplain}

\subsection{Double Debugger}
\label{subsec:double-debugger}

We call a \emph{double debugger} an interpreter that is able to run step by step
and display the program's state and current location in source, as well as the
state and location in the source of the interpreter itself. JSExplain is a
double debugger for JavaScript. The aim of this tool is to help a user to better
understand JavaScript's semantics, by providing an interpreter that is very
close to the specification of JavaScript. JSExplain's interface highlights the
expression of the program currently evaluated as well as the instruction of the
interpreter that is executed.

\subsection{Architecture}

The core of JSExplain is a JavaScript interpreter written in a subset of OCaml.
This interpreter is compiled to a subset of JavaScript and instrumented to
generate a trace of its execution when run. This trace can then be navigated
using a web-based
tool.\footnote{https://jscert.github.io/jsexplain/branch/master/driver.html}

The subset of OCaml we support for writing the interpreter is purely functional
with variables, constants, sequence, conditional, let-binding, function
definition, function application (with support for prefix and infix functions),
data constructors, records (including record projections, and the
``record-with'' construct to build a copy of a record with a number of fields
updated), tuples (i.e., anonymous records), and simple pattern matching (only
with non-nested patterns, restricted to data constructors, constants, variables,
and wildcards). For convenience, let-bindings and functions may bind simple
patterns (as opposed to only variables). We also support \texttt{ppx} extensions
for monadic programming, to simplify the handling of errors.

The target language of our compiler is a purely functional subset of JavaScript
where there is never any type conversion, where objects are used as records (no
use of their \texttt{prototype} field), and where tuples are encoded into
arrays.

ALAN: stopped here
